---
title: "2SLS"
author: "Denis Maciel"
date: "7/13/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
suppressPackageStartupMessages(library(tidyverse))

source(here::here("code", "load_datasets.R"))

load_datasets()
```


Prepare the data:

* Make treatment a numeric variable (0 for control, 1 for treatment)
* Extract the number of household members present in the lottery.
* Make any visit to emergency department binary and numeric

`ohp_all_ever_firstn_30sep2009`: This variable takes a value of 1 if an individual was enrolled in any Medicaid program (including the lotteried program, OHP Standard) between the earliest notification date in the sample (10 March 2008) and 30 September 2009. In the analysis of the 12-month mail survey data in Finkelstein et al (2012), this variable was used as the definition of insurance coverage in estimating the effect of Medicaid. In the analysis in Taubman et al (201XX), this variable was used as the definition of insurance coverage in estimating the effect of Medicaid.

```{r}
# Number of household in the list
eme_visits <- emergency %>% 
  left_join(descriptive %>% select(person_id, treatment, numhh_list)) %>% 
  left_join(state_programs %>% select(person_id, medicaid = ohp_all_ever_firstn_30sep2009)) %>% 
  mutate(n_household = dplyr::case_when(
    numhh_list == "signed self up" ~ 0L,
    numhh_list == "signed self up + 1 additional person" ~ 1L,
    numhh_list == "signed self up + 2 additional people" ~ 2L
  ))

# Treatment binary and numeric
eme_visits <- eme_visits %>% 
  mutate(treat_numeric = case_when(
    treatment == "Selected" ~ 1L,
    treatment == "Not selected" ~ 0L
  ))

eme_visits <- eme_visits %>% 
  mutate(any_visit_ed = case_when(
    any_visit_ed == "Yes" ~ 1L,
    any_visit_ed == "No" ~ 0L
  ))

eme_visits <- eme_visits %>% 
  mutate(medicaid = case_when(
    medicaid == "Enrolled" ~ 1L,
    medicaid == "NOT enrolled" ~ 0L
  ))

stopifnot(
  all(!is.na(eme_visits$n_household))
)

stopifnot(
  all(!is.na(eme_visits$treat_numeric))
)

stopifnot(
  all(!is.na(eme_visits$medicaid))
)

```

```{r}
eme_visits %>% 
  lm(data = ., formula = num_visit_cens_ed ~ treat_numeric + n_household)

eme_visits %>% 
  group_by(treatment) %>% 
  summarise(mean(num_visit_cens_ed, na.rm = TRUE))
```

## 2SLS

First Stage Regression: Regress Medicaid enrollment on the lottery

```{r}
first_stage <- lm(data = eme_visits, formula = medicaid ~ treat_numeric)

# add fitted values to dataframe
eme_visits$medicaid_hat <- first_stage$fitted.values
```

Second Stage Regression

```{r}
eme_visits %>% 
  lm(data = ., formula = num_visit_cens_ed ~ medicaid_hat + n_household)

eme_visits %>% 
  lm(data = ., formula = any_visit_ed ~ medicaid_hat + n_household)
```
