---
title: "Causal Forest"
author: "Denis Maciel"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)

source(here::here("code", "load_datasets.R"))
load_datasets()
```

```{r}
df <- emergency %>% 
  left_join(descriptive) %>% 
  left_join(state_programs) %>% 
  left_join(survey0) %>% 
  mutate(medicaid = ohp_all_ever_firstn_30sep2009) # varibel used by Taubman as Medicaid
```

## Feature Selection

Remove variables measured after the lottery. We want to be able to predict emergency department visits only with information before the lottery.

```{r}
ind <- which(
  !stringr::str_detect(colnames(df), "_ed") |
    stringr::str_detect(colnames(df), "pre_ed") |
    stringr::str_detect(colnames(df), "num_visit_cens_ed")
) 

df <- df[ , ind]
```

Remove columns with more than 5% NAs.

```{r}
ind <- which(
  map_dbl(df, function(col) sum(is.na(col)) / nrow(df)) < 0.05
)

df <- df[ , ind]
```

Calculate # of unique values and type of columms

```{r}
col_stats <- df %>% 
  map(function(col) 
    tibble(length = length(unique(col)), type = class(col))
  ) %>%
  reduce(bind_rows) %>% 
  mutate(col = colnames(df))

col_stats
```

Unclear meaning

```{r}
df <- df %>% 
  select(-starts_with("tanf_"),
         -starts_with("snap_"))

# If the respondents ever participated in OHP before 2009, i.e. after lottery
df <- df %>% 
  select(-starts_with("ohp_"))
```


```{r}
to_rmv <- c(
  "household_id", 
  "person_id",
  "draw_lottery", # The draw (from 1 to 8) one was picked should not be relevant
  "dt_retro_coverage", # Date where the coverage retroactively begins (probably coincides with lottery draw)
  "dt_notify_lottery", # Notification date about lottery 
  "postn_death", # Unclear 
  "zip_msa_list" # Has only one value across whole dataset
)

to_keep <- setdiff(colnames(df), to_rmv)

df <- df[, to_keep]
```

```{r}
df <- df %>% 
  mutate(age_2009 = 2009 - birthyear_list) %>% 
  select(-birthyear_list)
```

## Drops all NAs

```{r}
df <- drop_na(df)
```

## Mutate Variables

```{r}
# Number of household in the list
df <- df %>% 
  mutate(numhh_list = dplyr::case_when(
    numhh_list == "signed self up" ~ 0L,
    numhh_list == "signed self up + 1 additional person" ~ 1L,
    numhh_list == "signed self up + 2 additional people" ~ 2L
  ))

# Treatment binary and numeric
df <- df %>% 
  mutate(treatment = case_when(
    treatment == "Selected" ~ 1L,
    treatment == "Not selected" ~ 0L
  ))

df <- df %>% 
  mutate(medicaid = case_when(
    medicaid == "Enrolled" ~ 1L,
    medicaid == "NOT enrolled" ~ 0L
  ))

stopifnot(
  all(!is.na(df$numhh_list))
)
stopifnot(
  all(!is.na(df$medicaid))
)
stopifnot(
  all(!is.na(df$treatment))
)
```

Train vs test and preprocesing

```{r}
TRAIN_PROPORTION <- 0.3
ind <- sample(1:nrow(df), round(TRAIN_PROPORTION*nrow(df)), replace = FALSE)

train <- df[ind, ]
test <- df[-ind, ]

library(recipes)
rec <- recipe(num_visit_cens_ed ~ ., data = train %>% select(-medicaid, -treatment)) %>% 
  step_dummy(all_nominal())

prepped_rec <- prep(rec, train)

X_train <- juice(prepped_rec)
X_test <- bake(prepped_rec, new_data = test)
```

## Fit Model

```{r}
library(grf)

cf <- instrumental_forest(
  X = X_train,
  Y = train$num_visit_cens_ed,
  W = train$medicaid,
  Z = train$treatment,
  honesty = TRUE
)

# tree <- get_tree(cf, 1)
```

Predict on test set

```{r}
pred_df <- predict(cf,
                   newdata = X_test,
                   estimate.variance = TRUE)
```

Estimated treatment effects ranked

```{r}
pred_df %>% 
  arrange(predictions) %>% 
  mutate(idx = row_number()) %>% 
  ggplot(aes(idx, predictions)) +
  geom_point() +
  labs(x = "Rank",
       y = "Estimated Treatment Effect") +
  expand_limits(y = c(-5, 5)) +
  theme_light()
```

Estimated treatment effect with confidence intervals

```{r}
plot_htes <- function(cf_preds, z = 1.96) {
  ggplot(cf_preds, aes(rank(cf_preds$predictions), cf_preds$predictions)) +
    geom_point() +
    labs(x = "Rank", y = "Estimated Treatment Effect") +
    theme_light() +
    geom_errorbar(aes(
      ymin = cf_preds$predictions + z * sqrt(cf_preds$variance.estimates),
      ymax = cf_preds$predictions - z * sqrt(cf_preds$variance.estimates))
    )
}

# pred_df %>% 
#   # sample_frac(size = 1/100) %>% 
#   plot_htes(cf_preds =., ci = TRUE)

pred_df_sampled <- sample_n(pred_df, 200)
plot_htes(pred_df_sampled)
```

```{r}
pred_df %>% 
  bind_cols(test) %>% 
  mutate(bins = cut(predictions, breaks = seq(-5, 5, 0.5))) %>% 
  count(bins, medicaid) %>% 
  ggplot(aes(bins, n)) +
  # ggplot(aes(bins, n, fill = factor(medicaid))) +
  geom_col() +
  labs(x = "Bins", y = "") +
  theme_light() 
```

Feature importance

```{r}
feat_importance <- variable_importance(cf) %>% 
  as_tibble() %>% 
  rename(importance = V1) %>% 
  mutate(feature = colnames(X_train)) %>% 
  arrange(desc(importance))

feat_importance
```

